---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
### Upzip activity.zip 

``` {r upzip, message = FALSE}
unzip("activity.zip")

```

### Load and preview data
``` {r loadData}
act <- read.csv("activity.csv")
summary(act)
```

### Convert date into POSIXct class (ymd format)
```{r date_conversion, message=FALSE}
require(lubridate)
act$date <- ymd(act$date)
str(act)
```

## What is mean total number of steps taken per day?

### Calculate total number of steps taken per day
``` {r total_per_day, message = FALSE}
by_date_total <- data.frame(tapply(act$steps, act$date, sum, na.rm = TRUE))
colnames(by_date_total) <- "total_steps"
by_date_total$Date <- rownames(by_date_total)
```

### Make histogram of the total number of steps taken per day and report mean and median
``` {r hist, message = FALSE}
require(ggplot2)
ggplot(data=by_date_total, aes(x=as.factor(Date), y=total_steps))+
  geom_bar(stat = "identity")+xlab("Date")+ylab("Total Steps Per Day")+
  theme_bw()+theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 9))
```

### Calculate mean and median
```{r summary}
m_step <- mean(by_date_total$total_steps, na.rm = TRUE)
med_step <- median(by_date_total$total_steps, na.rm = TRUE)
```
The mean total number of steps taken per day is `r round(m_step,2)` steps, and 
the median total number of steps take per day is `r med_step` steps.

## What is the average daily activity pattern?

### Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r pattern, message=FALSE}
require(dplyr)
daily_act <- act%>%group_by(interval)%>%summarise(mean=mean(steps,na.rm=TRUE))

ggplot(data=daily_act, aes(x=interval, y=mean))+
  geom_line()+xlab("Intervals")+
  ylab("Average number of steps taken across all day")+
  theme_bw()
```

### Find the 5-minute interval that contains the maximum number of steps
```{r step_max}
step_max <- daily_act$interval[which.max(daily_act$mean)]
```
The interval `r step_max` contains the maximum number of steps of `r round(max(daily_act$mean),2)`

## Imputing missing values

### Calculate the number of NAs in the dataset
```{r na}
numNA <- sum(is.na(act$steps))
```
The number of NAs in the dataset is `r numNA`

### Devise strategy to fill in the missing values in the dataset
The missing values will be replaced with the median of the corresponding 5-min interval.

### Create new dataset with missing values filled in
```{r full}
replace <- daily_act[match(filter(act,is.na(steps))$interval,daily_act$interval),"mean"]
act[is.na(act$steps),"steps"] <- as.vector(replace)
```

### Make histogram of the total number of steps taken per day and report mean and median
``` {r full_hist, message = FALSE}
by_date <- data.frame(tapply(act$steps, act$date, sum, na.rm = TRUE))
colnames(by_date) <- "total_steps"
by_date$Date <- rownames(by_date)

ggplot(data=by_date, aes(x=as.factor(Date), y=total_steps))+
  geom_bar(stat = "identity")+xlab("Date")+ylab("Total Steps Per Day")+
  theme_bw()+theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 9))

mean_step <- mean(by_date$total_steps, na.rm = TRUE)
median_step <- median(by_date$total_steps, na.rm = TRUE)
```
The mean total number of steps taken per day is `r round(mean_step,2)` steps, and 
the median total number of steps take per day is `r median_step` steps.
Both mean and median increase with the input of missing values.

## Are there differences in activity patterns between weekdays and weekends?

### Make an indicator for whether a given day is weekday or not
```{r weekday}
act$isWeekDay <- ifelse(weekdays(act$date)=="Sunday" | weekdays(act$date)=="Saturday",
                        "Weekend", "Weekday")
```

### Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days, comparing weekday and weekend
```{r weekPlot}
week_act <- data.frame(tapply(act$steps, list(act$interval, act$isWeekDay), 
                              mean, na.rm = TRUE))
week_act$Interval <- rownames(week_act)

require(reshape)
act_plot <- melt(week_act)
ggplot(data=act_plot, aes(x=as.numeric(Interval), y=value))+
    geom_line()+xlab("Intervals")+
  ylab("Average number of steps taken across all day")+
  theme_bw()+facet_wrap(.~variable)
```